Sketch Recognition and Understanding

### 2009-09-01 draft/System.tex
after dr. mahmoud meeting

 %the performance of most algorithms is based on user style.

% Stroke segmentation is one of the most complex problems in the sketch recognition system. It is dividing the   
%In stroke segmentation problem is Particle swarm optimization is used as the

\section{System Overview}
\label{sec:AnOverviewOfTheSystem}
   This research solves sketch segmentation problem using particle swarm algorithm (PSO).  Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information. After that a set of possible corner or critical points is computed to guide the segmentation procedure.  A set of features is extracted from the segmented strokes to be used by the classifier.  %The stroke is the path of points from the instant the pen is down till it is up.
The block diagram in Figure \ref{fig:Blockdiagram} shows the main system blocks. The next section describes each block in details.
 %Due to sloppiness of users and hardware glitches the system processes the data before further
\begin{figure}[]
	\centering

\begin{center}
	\includegraphics[scale=0.75]{images/AllBlockDiagram.eps}
	\caption[The System Block Diagram]{Block diagram of the system}
	\label{fig:Blockdiagram}
\end{center}
\end{figure}

@@

The next step is \textbf{Segmentation}, the goal of the segmentation stage is to divide strokes into segments of either curves or lines. The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive. This is an optimization problem, which evolutionary programming can solve efficiently. A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.

 As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. If the stroke proved to be an elliptical arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either a polygon or a set of lines and curves. The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%
% First an attempts the segmentation process is divided into two steps ellipse fitting and curve segmentation. In the first step, the ellipse fitting process tries to fit the stroke into an ellipse. If the system fails it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation. % is the ellipse fitting

@@

 In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes. The feature vector is used as input for the next step.   %the system computes composite set of features some are statistical other are spatial features based on the type primitives.% The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.

   The final step is \textbf{Classification} using a SVM classifier. The classifier attempts to classify the symbol into one the previously trained classes.% c  that will use the features computed to classify the segments into one of the previously trained classes. % Or to determine the symbol of the given segments from the set of preciously trained symbols.

 %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives. Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\

%\chapter{System Details}
%\label{sec:SystemDetails}

\section{Possible Dominant Point Extraction}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The Possible Dominant Point Extraction step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

@@

\section{Segmentation}
\label{sec:Segmentation}
The segmentation tries to divide the stroke into a set of primitives. As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves. The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves. The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
%If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves (see the block diagram in fig. \ref{fig:Blockdiagram} ). 
%The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig. \ref{fig:Blockdiagram}).  The next sections will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.
%After computing the primary data the system now tries \\
%Paragraphs describe the segmentation algorithm.

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.

@@

\subsection {Training and Testing}
A training set that contains positive samples of each category in the data set is used to train the SVM classifier.  The training starts after computing the feature vector of each sample. The system trains $ \frac{n(n-1)}{{2}}$ binary classifiers each distinguish between only two categories. The goal of the training process is to generate a linear SVM plane that maximize margin between the two categories.  A cross validation procedure is implemented to choose the best value for the $C$ paramter in SVM classifier.( $C=100$ was the best value) 
=
For classification the decision is made from  producing the feature vector of the unkown symbol to the trained $\frac{n(n-1)}{{2}}$ binary classifiers. The final decision is given using simple voting strategy.

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gaining increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors. The generated plane is used as a good classifier to the unseen examples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calculate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier planes }
	\label{fig:svm1}
\end{figure}

@@

 The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.   Lets make an assumption that the datat point are linearly separable, i.e. there are infinite possibilities to find a plane that separates and correctly classify the training data. This is idea is illustrated in Figure\ref{fig:svm1} where two different separating planes can correctly classify the data. It is intuitively noted from the figure that the solid line is more likely to generalize more data. This means less classification error will be introduced. The geometrical characterization of this plane means the plane furthest from both classes.     % Which one is preferable ? Intuitively one perfers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furthuret from both classes.  plane supports a class if all points in that class are on one side of that plane.

@@

The Support vectors that are pumped into the supporting plane determine the margin. This is a primal form Quadratic Programming (QP) problem. The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximumm margin between parallel 
\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
  \end{array} 
\label{eq:dual}
\end{equation}

@@

The generalization error on future points and not only in the training set can be obtained as proved by the statical learning theory \cite{BennettSVMP2000}.  The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexity or capacity of the classification. In linear functions, using planes to maximize the margin of separation reduces the complexity of the functions. Thus, by explicitly maximizing the margin we are in the same time  minimizing bound on the generalization error and therefore expect better generalization with high probability. The dimensionality of the data dose not affect the size of the margin which   results in good performance  for very high dimensional data (i.e, with a very large number of attribute). In a sense, this also reduces the  problems caused by over fitting of high - dimensional data.  A large volume of literature  , e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theory.

%Using geomteric arguments we can gain in sights to some of these results. Classification function that have capcity to fit the training data are more likely to overfit  resulting in poor generalization.

\subsubsection{Linearly Inseparable Case}
In all previous sections we assumed that the problem of classifying two set is linearly separable. The assumption is not true in most cases, thus, trying to construct a plane to separate the two dataset will fail. Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly. The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.  Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin. But we must relax the constrain that insure that each point in on the appropriate side of tis supporting plane. Every point that falls on the wrong side of the supporting plane is considered an error. The problem is now to maximize the margin and minimize the error.

@@

\subsubsection {Non linear function Using Kernels}
Figure \ref{fig:svm6} shows a classification problem that cannot be separated using linear space. A higher order plane can work better in similar cases. Converting a linear classification algorithm into non linear classification algorithm  can be achieved by adding additional attribute to the data that are nonlinear functions of the original data. The expanded dataset features space can be separated using an  existing linear classification algorithm.  The  linear classification that is applied on the feature space produce a non linear classification on the original input space. In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminant  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminant in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

For higher dimensional datasets, this method of nonlinear mapping  has two potential problems stemming from the fact that dimensionality of the feature space explones exponentially. The first problem is that over fitting becomes a problem. However, since SVM rely on margin maximization they are largely immune to this problem. Provided an appropriate value of parameter $C$ is chosen. The second concern is that it is the computational complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.

Examine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.  By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernel are being developed to fit domain specific requirements. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
		Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\

	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product. Thus by changing kernels we can get different highly nonlinear classifiers. No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product. All the benefits of the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

%The foundations of Support Vector Machines (SVM) have
\subsubsection{Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Select parameter C representing the tradeoff between minimizing the training set error and maximizing the margin. Select the kernel function and any kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
	\item Recover th primal threshold variable $b$ using the support vectors. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}

The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
\subsubsection {Multi Class Classifiers}

 There is mainly two popular methods to change the binary SVM classification problem into Muliclass classification problem. The first methods is One Versus All (OVA) Strategy where M binary classsifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negtive example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.

>>>>

%the performance of most algorithms is based on user style. 
% Stroke segmentation is one of the most complex problems in the sketch recognition system. It is dividing the   
%In stroke segmentation problem is Particle swarm optimization is used as the 
\section{System Overview}
\label{sec:AnOverviewOfTheSystem}
   This research solves sketch segmentation problem using particle swarm algorithm (PSO).  Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information. After that a set of possible corner or critical points is computed to guide the segmentation procedure.  A set of features is extracted from the segmented strokes to be used by the classifier.  %The stroke is the path of points from the instant the pen is down till it is up.
The block diagram in Figure \ref{fig:Blockdiagram} shows the main system blocks. The next section describes each block in details.
 %Due to sloppiness of users and hardware glitches the system processes the data before further
\begin{figure}[]
	\centering

\begin{center}
	\includegraphics[scale=0.9]{images/AllBlockDiagram.eps}
	\caption[The System Block Diagram]{Block diagram of the system}
	\label{fig:Blockdiagram}
\end{center}
\end{figure}

@@

The next step is \textbf{Segmentation}, the goal of the segmentation stage is to divide strokes into segments of either curves or lines. The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive. This is an optimization problem, which evolutionary programming can solve efficiently. A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve. However, comparisons between PSO and GA \cite{ComparePSOGA05} shows that PSO converges to optimal solution faster than GA.

 As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}. If the stroke proved to be an elliptical arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either a polygon or a set of lines and curves. The algorithms takes the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%

% First an attempts the segmentation process is divided into two steps ellipse fitting and curve segmentation. In the first step, the ellipse fitting process tries to fit the stroke into an ellipse. If the system fails it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms. The two algorithms will generate two segmentations, the system will choose the segmentation that has the minimum segmentation error.%  generate the segmentations,  the minimum error will be the chosen segmentation. % is the ellipse fitting

@@

 In \textbf{Feature Extraction} step, the approximated segments that were computed in the previous step are appended to a list of segments. A feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes. The feature vector is used as input for the next step.   %the system computes composite set of features some are statistical other are spatial features based on the type primitives.% The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.

   The final step is \textbf{Classification} using a SVM classifier. The classifier attempts to classify the symbol into one the previously trained classes.
   % c  that will use the features computed to classify the segments into one of the previously trained classes. % Or to determine the symbol of the given segments from the set of preciously trained symbols. 
 %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives. Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
%\chapter{System Details}
%\label{sec:SystemDetails}
\section{Possible Dominant Point Extraction}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The Possible Dominant Point Extraction step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

@@

\section{Segmentation}
\label{sec:Segmentation}
The segmentation tries to divide the stroke into a set of primitives. As shown in Figure \ref{fig:Blockdiagram} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov}. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves. The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed during Possible Dominant Point Extraction then produce a set of dominant points which are connected by either lines or curves. The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.
%If the stroke proved to be an ellipse then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed into two particle swarm algorithms that will divide the stroke to either lines or lines and curves (see the block diagram in fig. \ref{fig:Blockdiagram} ). 
%The algorithms takes the stroke points along with the possible dominant points computed then produce a set of dominant points which are connected with either lines or curves (see fig. \ref{fig:Blockdiagram}).  The next sections will describe the ellipse detection algorithm and both the two particle swarm algorithms used to divide the stroke.
%After computing the primary data the system now tries \\
%Paragraphs describe the segmentation algorithm.

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold which is computed as 20 \% of the input stroke area.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.

@@

\subsection {Training and Testing}
A training set that contains positive samples of each category in the data set is used to train the SVM classifier.  The training starts after computing the feature vector of each sample. The system trains $ \frac{n(n-1)}{{2}}$ binary classifiers each distinguish between only two categories. The goal of the training process is to generate a linear SVM plane that maximize margin between the two categories.  A cross validation procedure is implemented to choose the best value for the $C$ parameter in SVM classifier.( $C=100$ was the best value) 
=
For classification the decision is made from producing the feature vector of the unknown symbol to the trained $\frac{n(n-1)}{{2}}$ binary classifiers. The final decision is given using simple voting strategy.

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gaining increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors. The generated plane is used as a good classifier to the unseen examples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier planes }
	\label{fig:svm1}
\end{figure}

@@

 The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.   Let's make an assumption that the data point are linearly separable, i.e. there are infinite possibilities to find a plane that separates and correctly classify the training data. This is idea is illustrated in Figure\ref{fig:svm1} where two different separating planes can correctly classify the data. It is intuitively noted from the figure that the solid line is more likely to generalize more data. This means less classification error will be introduced. The geometrical characterization of this plane means the plane furthest from both classes.     % Which one is preferable ? Intuitively one prefers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furtherer from both classes.  Plane supports a class if all points in that class are on one side of that plane.

@@

The Support vectors that are pumped into the supporting plane determine the margin. This is a primal form Quadratic Programming (QP) problem. The concept of duality can also used to wields the problem to a Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximum margin between parallel 
\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
  \end{array} 
\label{eq:dual}
\end{equation}

@@

The generalization error on future points and not only in the training set can be obtained as proved by the statistical learning theory \cite{BennettSVMP2000}.  The generalized error bound is a function of the misclassified error on the training data and the terms that measure the complexity or capacity of the classification. In linear functions, using planes to maximize the margin of separation reduces the complexity of the functions. Thus, by explicitly maximizing the margin we are in the same time minimizing bound on the generalization error and therefore expect better generalization with high probability. The dimensionality of the data does not affect the size of the margin which   results in good performance for very high dimensional data (i.e, with a very large number of attribute). In a sense, this also reduces the problems caused by over fitting of high - dimensional data.  A large volume of literature, e.g., \cite{bookSVMoverfit11,statisticalSVM1} refers to this problem,  for more technical discussion of the   statistical theory.

%Using geometric arguments we can gain insights to some of these results. Classification function that have capacity to fit the training data are more likely to over fit resulting in poor generalization.

\subsubsection{Linearly Inseparable Case}
In all previous sections we assumed that the problem of classifying two set is linearly separable. The assumption is not true in most cases, thus, trying to construct a plane to separate the two dataset will fail. Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly. The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.  Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin. But we must relax the constrain that insure that each point in on the appropriate side of it is supporting plane. Every point that falls on the wrong side of the supporting plane is considered an error. The problem is now to maximize the margin and minimize the error.

@@

\subsubsection {Non linear function Using Kernels}
Figure \ref{fig:svm6} shows a classification problem that cannot be separated using linear space. A higher order plane can work better in similar cases. Converting a linear classification algorithm into non linear classification algorithm can be achieved by adding additional attribute to the data that are nonlinear functions of the original data. The expanded dataset features space can be separated using an existing linear classification algorithm.  The linear classification that is applied on the feature space produces a non linear classification on the original input space. In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminate  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminate in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

For higher dimensional datasets, this method of nonlinear mapping has two potential problems stemming from the fact that dimensionality of the feature space explodes exponentially. The first problem is that over fitting becomes a problem. However, since SVM rely on margin maximization they are largely immune to this problem. Provided an appropriate value of parameter $C$ is chosen. The second concern is that it is the computational complexity of   computing $\vartheta(x)$ but SVM use kernels to get around this problem.

Examine what happens when  the nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define: $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.  By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernels are being developed to fit domain specific requirements. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product. Thus by changing kernels we can get different highly nonlinear classifiers. No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product. All the benefits of the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

%The foundations of Support Vector Machines (SVM) have
\subsubsection{SVM  Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Selecting the SVM classifier Paramters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
	\item Get the support vector and recover the emprial threshold  variable $b$. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}
The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
\subsubsection {Multi Class Classifiers}

 There is mainly two popular methods to change the binary SVM classification problem into Multiclass classification problem. The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.
